NOTE THIS CONTENT WAS MOVED TO THE INDEX AND THE README FILE

# Contributing to this guide

We need your help.  Please contribute or insert comments and your experiences in this guide.  This is written in a form of [RMarkdown](https://r4ds.had.co.nz/r-markdown.html) called "bookdown."  (there will evenntually be a chapter here about Rmarkdown)

Please contribute if you've never used bookdown.  We are delighted if you want to contribute and don't want you to worry about formatting or bookdown tricks.   We are interested in your ideas.   Simply edit any one of the chapter files (Rmd files that begin with a two digit numeral) as you would edit an Rmarkdown file.   If you want to add a new chapter, simply start a new Rmarkdown file, preface with a header  (e.g. `# Tracking Woozles`) which will be the chapter title.   We can incorporate these changes and edit for bookdown tech as needed.   

This project exists on a public github project, so please fork, make changes and create a merge request.  

Alternatively if you are unsure or just have a short idea or correction, you may submit a github issue to this project (see [Github.com:Creating an issue](https://help.github.com/en/github/managing-your-work-on-github/creating-an-issue) which requires a github account.   

If you are a lab collaborator and have a found a successful technique we want to hear about it.   Consider a new chapter describing this and pointers to examples on github.  

## Prerequisites

You need a plain text editor that can create Markdown (we use Rstudio).  Word Processors insert unreadable characters (curly quotes).  When writing markdown ou can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

To render this book, you need the **bookdown** package, which can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

it also may require the pandoc utility.  On mac open the terminal install the (homebrew)[https://brew.sh/] utility using there instructions,  and then  `brew install pandoc`


Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this book to HTML for github the command is ` bookdown::render_book("index.Rmd", "bookdown::gitbook", output_dir = "docs")`   Alternatively you may test the book using  `bookdown::serve_book(output_dir = 'docs')` which renders, then starts a webserver to preview it.  The project is set to render the HTML into the 'docs' folder which will make it easy to publish as an HTML website (e.g. github pages), but the "output_dir" param makes it explicit. 

To compile this book to to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
